### 2020-05-04 birmingham.Rmd

# Birmingham

```{r birmingham}
# idea: use buffers of LSOA centroids to supplent built-up areas
# birmingham_centroids = get_pct_centroids("west-midlands", geography = "lsoa")
# birmingham_lsoa_buff = stplanr::geo_buffer(birmingham_centroids, dist = 1000)
# mapview::mapview(birmingham_lsoa_buff)
# birmingham_res_buff = st_intersection(builtup, birmingham_lsoa_buff) %>% st_union()
# roads = roads %>%
#   st_transform(27700) %>%
#   st_intersection(birmingham_res_buff) %>%
#   st_transform(4326)
# idea: look for other basemaps
region = las_top %>% filter(name == "Birmingham")
roads = rsf %>% filter(region == "Birmingham", pctgov > 100)
roads = st_intersection(roads, region) # remove roads outside region

roads_top = roads %>%
  filter(length > 300) %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) + tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```