### 2020-05-04 cambridge.Rmd

# Cambridge

```{r cambridge}
roads = rsf %>% filter(region == "Cambridge", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Cambridge", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 3000)
roads = st_intersection(roads, region_buff)

roads_top = roads %>%
  filter(length > 100) %>%
  filter(name != "") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 13) + tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```