### 2020-05-06 london.Rmd

# London

```{r london}
region = las_top %>% filter(name == "London")
roads = rsf %>% filter(region == "London", pctgov > 500)
# roads = roads %>% st_transform(27700) %>% st_intersection(london_res_buff) %>% st_transform(4326)

roads_top = roads %>%
  filter(length > 500) %>%
  filter(!name == "Talgarth Road", length > 500) %>%
  filter(!name == "Whitechapel Road", length > 500) %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
# tm_shape(roads_top) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
#   tm_text("name") +
#   tm_view(set.view = 12) + tm_scale_bar()
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = 12) +
  tm_scale_bar() +
  tm_basemap(server = s)
roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```