### 2020-05-07 article.Rmd
Article template

---
title: "Methods to support cycleways and pedestrianised spaces in response to the Covid-19 pandemic: overview and findings from 10 UK cities"
author: "Robin Lovelace"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- Guidance from https://transportfindings.org/for-authors -->

<!-- Transport Findings welcomes research findings in the broad field of transport. Articles must either pose a New Question,  present a New Method, employ New Data (including New Contexts or Locations),  discover a New Finding (i.e. it can almost exactly replicate a previous study and find something different), or some combination of the above. -->
<!-- Scope -->

<!-- You may find yourself asking if your paper is within the scope of Transport Findings. -->

<!--     Is there a hypothesis somehow related to transport? -->
<!--     Is there a (scientifically valid, replicable) methodology? -->
<!--     Is there a finding? -->

<!-- If you can answer yes to these questions, it is within scope. -->
<!-- Article Types -->

<!--     Findings - where the object of study is nature -->
<!--     Syntheses - where the object of study is the literature -->
<!--     Cases - where the objects of study are particular sites or projects, and methods may be more qualitative -->

<!-- Sections -->

<!-- All articles shall have 3 sections, and only 3 sections, titled as follows: -->

<!--     RESEARCH QUESTION[S] AND HYPOTHESIS[ES] -->
<!--     METHODS AND DATA -->
<!--     FINDINGS -->

<!-- There shall be no introduction, “road-map paragraph,” literature review, conclusions, speculations, or  policy implications beyond what is included above. Focus on what you found, not why you found it. -->
<!-- Submissions -->

<!-- The manuscript submission must include the following: -->
<!-- TITLE -->
<!-- AUTHORS (NAME, AFFILIATION, CONTACT) -->
<!-- ABSTRACT -->
<!-- KEYWORDS -->
<!-- ARTICLE (Sections 1, 2, 3) -->
<!-- ACKNOWLEDGMENTS -->
<!-- REFERENCES -->

<!-- Manuscript submissions may include SUPPLEMENTAL INFORMATION in separate files that do not count against article length. This information should not be essential for the general understanding of the manuscript. -->
<!-- Style -->
<!-- Focus and Parsimony -->

<!-- Papers should be focused and to the point, and not begin with trite observations like “Congestion is a problem the world over.” Usually you can delete your opening paragraph if it begins like that, and the reader is no worse off. As Strunk and White say: “Omit Needless Words”. The Abstract should not say the same thing as the Introduction. -->
<!-- Transparency and Replicability -->

<!-- A minimum standard for a good paper is transparency and replicability: Can the reader understand what you did, and repeat it, and get the same answer? -->
<!-- Mathematical Conventions -->

<!-- Each variable shall have one, and only one, definition per document. -->

<!-- Each defined term in the document shall be represented by one and only one variable. -->

<!-- Lowercase and uppercase versions of the same letter should be logically related. For instance, use lowercase letters to define the PDF (probability distribution function) or individual instance, and uppercase letters the CDF (cumulative distribution function) or population, so when you sum:  i=1 to I, k=1 to K, etc. -->

<!-- All variables shall be a single letter or symbol. Double or triple letter variables can be confused with multiplication. If you have more than 52 symbols in your paper (26 letters for both lower and upper case), consider (a) there are too many, and (b) using Greek or Hebrew characters. -->

<!-- Use subscripts liberally to differentiate things that, for instance, are of a class but measured differently, or computed with different assumptions. -->

<!-- All equations shall have all of their variables defined. -->