### 2020-05-09 cities/bristol.Rmd
Tidy-up

# Bristol

```{r bristol}
region = las_top %>% filter(name == "Bristol")
region = sf::st_make_valid(region)
roads = rsf %>% filter(region == "Bristol", pctgov > 100)
roads = st_intersection(roads, region)

bristol_centre = tmaptools::geocode_OSM("Bristol", as.sf = TRUE)
bristol_buff = stplanr::geo_buffer(bristol_centre, dist = 1500)
roads_top = roads[bristol_buff, ] %>%
  filter(length > 300) %>%
  filter(name != "Temple Way Underpass") %>%
  filter(!name == "Newfoundland Street", !name == "") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)

tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-2.5879, 51.454, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```