### 2020-05-09 cities/leeds.Rmd

# Leeds

```{r leeds}
leeds_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
leeds_buff = stplanr::geo_buffer(leeds_centre, dist = 4000)
roads = rsf %>% filter(region == "Leeds", pctgov > 100)
region_centre = tmaptools::geocode_OSM("Leeds", as.sf = TRUE)
region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
roads = st_intersection(roads, region_buff)

roads_top = roads[leeds_buff, ] %>%
  filter(length > 300) %>%
  filter(name != "Broadway") %>%
  filter(name != "Wellington Street") %>% 
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-1.549, 53.805, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```