### 2020-05-09 cities/liverpool.Rmd

# Liverpool

```{r liverpool}
roads = rsf %>% filter(region == "Liverpool", pctgov > 100)
region = las_top %>% filter(name == "Liverpool")
roads = st_intersection(roads, region)
# region_centre = tmaptools::geocode_OSM("Liverpool", as.sf = TRUE)
# region_buff = stplanr::geo_buffer(region_centre, dist = 9000)
# tm_shape(roads) + tm_lines("pctgov", lwd = 3, palette = "viridis") +
#   tm_view(set.view = 10) + tm_scale_bar() +
#   tm_shape(region) + tm_borders()

roads_top = roads %>%
  filter(length > 200) %>%
  filter(name != "Queensway (Mersey Tunnel)") %>%
  filter(name != "Queensway") %>%
  arrange(desc(pctgov)) %>%
  slice(1:10)
tm_shape(roads) + tm_lines("lightblue", lwd = 3, alpha = 0.7) +
  tm_shape(roads_top) + tm_lines(col = "blue", lwd = 3) +
  tm_text("name") +
  tm_view(set.view = c(-2.95, 53.4084, 13)) + 
  tm_scale_bar() +
  tm_basemap(server = s)

roads_top %>%
  st_drop_geometry() %>%
  select(name, ref, road_type = highway, cycling_potential = pctgov, length_m = length) %>%
  knitr::kable(digits = 0)
```