### 2020-05-11 conversation.Rmd
Tweak title + fix typos

---
title: "Pop-up cycleways: a peak into post COVID-19 transport futures"
---

Around the world, as nations prepare to emerge from months of COVID-19 induced lockdown, life cannot simply return to previous norms. The pandemic has transformed all sectors of the economy, not least transport. Long distance travel has been dramatically altered, as [demand plummets](https://osf.io/preprints/socarxiv/v3g5d/) and [airline companies go bust](https://arxiv.org/abs/2004.08460). Yet short distance travel patterns have also changed greatly. Forced to stay close to home, many people have started walking and cycling more. From a health and well-being perspective, this may be a more important shift, especially in the long term as lockdown measures are gradually eased.

Active travel is already [recommended by the World Health Organisation](https://who.canto.global/pdfviewer/viewer/viewer.html?share=share%2Calbum%2CU6GDM&column=document&id=q3v02qdsh114763ffe3apo4s2p&suffix=pdf) as a safe mode of transport, and in the coming weeks and months it will play a vital role in enabling the reopening of many countries. As travel increases, we must ensure that safe, socially distanced walking and cycling remains possible. Maintaining social distancing on public transport could severely reduce the capacity of rail and bus networks, and there is a risk that this may lead to a rapid rise in road traffic.  [ record bicycle sales ] suggest that many people have taken advantage of lockdown conditions to take up or return to cycling, but without careful management, increased traffic congestion could again deter many would-be cyclists from the streets.

Active travel has extensive [ social and environmental benefits ]. Physical activity is associated with XX health conditions….. mental health…..isolation...

Evidence is beginning to emerge of [ potential links ] between air pollution and COVID-19 lethality, not least because many health conditions such as asthma and xxx are [ worsened by poor air quality ]. Indeed, it is estimated that global improvements in air quality in recent months have led to [ xx thousand fewer deaths ] from conditions such as these. This makes it all the more important to restrict any rise in road traffic and to encourage more beneficial means of travel.

Local and national governments worldwide are still working out how best to respond to these challenges. On Saturday 9th May, British Transport Secretary Grant Shapps announced a £2 billion investment in measures to promote walking and cycling. This follows a £10 million commitment from the Scottish Government and action announced in Wales. Hundreds of kilometres of temporary infrastructure is being put in place in dozens of cities worldwide, from [Milan] to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists ""), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf). Never has there been such a need for affordable, healthy and socially distanced transport that does not create the fumes that [emerging evidence] links to coronavirus susceptibility.

In terms of practical steps to encourage walking and cycling, a range of simple interventions can be undertaken, initially on a temporary basis. Roads can be closed to motor vehicles, or physical barriers can be used to enable [ filtered permeability ], in which access for residents is retained but modal filters enable routes for walking and cycling that are faster and more direct than the routes for motor vehicles. This is well established in [ the London Borough of Hackney ] where attractive barriers such as planters have been used to good effect. Another option involves keeping the roads open while reallocating excess roadspace to widen pavements or introduce pop-up cycleways.

These measures can all be undertaken on a temporary basis, but if action is prioritised in places where there is a *long-term need* for cycle improvements, there is a greater chance that people will demand they become permanent. This is what happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. In Paris, measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).

With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. As with the medical science, research is needed now. Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed. Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen. Today the greatest need may be for medical workers to safely access [hospitals], tomorrow it could be access to schools or retail centres.

Building on the previous projects such as the [ Propensity to Cycle Tool ] and the [ Cycling Infrastructure Prioritisation Toolkit ], researchers from the University of Leeds have investigated where roadspace could be reallocated for pop-up cycleways. They have mapped out the top locations for intervention in ten major cities across England, with the intention of rolling out these methods across the country in the near future.

The methods have focused on road capacity and cycling potential, identifying roads where a ‘spare lane’ could be re-purposed for cycling. Cycle potential is assessed based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025. The length of road segments is also used to filter the top schemes.

The use of open data and publicly accessible models is a key aspect of this research. Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible. The potential for international application can be seen for example in [ WHO report ] modelling cycling uptake in cities in low income countries. Within the UK, [ 50+ Local Authorities ] are now using the PCT to inform their strategic cycle network plans.

Maps (see below) show the recommendations for where to introduce pop-up cycleways. With the situation rapidly developing, some of the roads shown here, such as [Princess Street in Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175), have already been identified by Local Authorities as targets for intervention.

<!-- # Where is pop-up infrastructure most needed? -->

<!-- # Story  -->

<!-- The COVID-19 pandemic has transformed all sectors of the economy. -->
<!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
<!-- Yet short-distance travel patterns have also changed greatly and, from a health and well-being perspective, this may be a more important shift. -->
<!-- With temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. -->
<!-- With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. -->
<!-- As with the medical science, research is needed now. -->
<!-- Methods developed to identify locations of high walking and cycling potential can help ensure that the 'pop-up' infrastructure that goes in now is effective, safe, and placed where it is most needed. -->

<!-- (this would be two to three sentences summarizing the article’s focus) -->

<!-- # Significance  -->

<!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> 
<!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post COVID-19 "new golden age for cycling". -->
<!-- Without evidence there is a risk that these words are not supported by effective and long-term interventions. -->
<!-- Our research into prioritising the right kind of cycling infrastructure where it is most needed can ensure that decisions made now are of maximum benefit for the coming months when physical distancing will need to continue. -->
<!-- Polls showing people do not want to 'return to normal' also suggest that the COVID-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
<!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. -->

<!-- # Key points -->

<!-- - Our analysis of data generated for the Department for Transport funded Propensity to Cycle Tool (PCT) and Cycling Infrastructure Prioritisation Toolkit (CyIPT) has already flagged dozens of roads in 10 major cities across England where a 'spare lane' could be replaced by low-cost bidirectional temporary infrastructure.  -->
<!-- - A method highlights roads based on number of lanes, cycling potential and length. -->
<!-- - Feedback from cycle campaigners coordinated by CyclingUK suggests the results for 10 cities (London, Birmingham, Manchester, Leeds, Liverpool, Newcastle, Sheffield, Cambridge, Bristol, Leicester) match intuition and can support local knowledge and community engagement in support of temporary infrastructure. -->
<!-- - Further work, which we are discussing with the Department for Transport, will develop more sophisticated methods to provide evidence that is urgently needed to support pop-up spaces in the UK and beyond. -->

<!-- CyclingUK are putting out a press release as an exclusive to national media organisations. -->
<!-- This article will provide a more detailed view to go out the following day while interest in the topic is at a peak. -->

<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post COVID-19 recovery package in Greater Manchester. -->

<!-- ![](https://raw.githubusercontent.com/cyipt/tempCycleways/master/README_files/figure-gfm/manchester-1.png) -->

<!-- Other key points(anything additional that you want to make sure they consider/know about) Highlighting the website further etc and we would also mention the BBC exclusive here. I’ve never had a problem with the Conversation agreeing to hold off on publishing an article until an embargo or exclusive deadline has been met. However it would be good to make clear how the Conversation article will differ from the BBC coverage.   -->

 <!-- supporting the case for reclaiming road space for safe, socially distanced key worker travel. -->