### 2020-05-12 conversation.Rmd
Update articles

---
title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
---

As nations prepare to emerge from months of COVID-19 induced lockdowns, it is becoming clear life cannot simply return to normal. The pandemic has transformed all sectors of the economy, not least transport.

In the coming weeks and months, walking and cycling will play a vital role in enabling the reopening of many countries, as continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity. There is a window of opportunity here that could be missed if roads become busy and dangerous again before new active transport habits become embedded.

Local and national governments are scrambling to act. In the UK, £2 billion investment in measures to promote walking and cycling has been announced. £250 million of this is for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020). New [guidance from the Department for Transport](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities/traffic-management-act-2004-network-management-in-response-to-covid-19) advocates "measures to reallocate road space to people walking and cycling".

<tc-read-more id="137909"></tc-read-more>

Local authorities urgently need answers on how this funding should be spent, and where road space reallocation should be prioritised. New cyclinginfrastructure which is developed based on rigorous analysis of the best available data, in combination with local knowledge and community engagement, are more likely to be effective than plans based on intuition alone. Recognising the importance of a strong and actionable evidence-base, and the window of opportunity, we used our experience developing tools that are already being used by local authorities, notably the multi-institution PCT and the CyIPT, to develop methods and outputs taylored for the new transport planning landscape.

<image id="334100" align="centre" caption="Screenshot from the website salfordliveablestreets.commonplace.is to support local responses to the COVID-19 pandemic, including the prioritisation of pop-up active transport infrastructure." />

In a pre-print [paper](https://osf.io/7wjb6) on methods to prioritise pop-up road space reallocation schemes, we have classified pop-up interventions into three categories. Firstly, in streets that are wide or have multiple lanes, roadspace can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625). Secondly, physical barriers can be used to make routes for walking and cycling [faster and more direct](https://cyclehighways.eu/design-and-build/infrastructure/filtered-permeability.html) than the routes for motor vehicles. This can be seen in Tower Hamlets, London, where simple barriers like planters are used to good effect. Thirdly, roads can be closed altogether to motor vehicles, as is happening in [Deansgate, Manchester](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175).

The first of these three options is the most simple to implement,  with relatively few implications for traffic flow. We have mapped out **link** the top locations for this reallocation of road space in ten major cities across England, with the intention of rolling out these methods across the country in the near future. Having identified roads where a spare lane could be re-purposed for cycling, we have assessed the cycle potential of these streets based on census data for travel to work and cycling uptake scenarios under the UK Government’s national target of doubling cycling levels by 2025. The length of road segments is also used to filter the top schemes.

The use of open data and publicly accessible models is a key aspect of this research. Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible. The potential for international application can be seen for example in [research for the World Health Organisation](https://atfutures.github.io/upthat/articles/adaptation.html) modelling cycling uptake in cities in low income countries. Within the UK, [50+ Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.

Results from our [methods](https://osf.io/7wjb6) show that many cities have 'spare lanes' under the current levels of traffic in city centres and near key worker destinations such as hospitals. These results could support the prioritisation of vital resources dedicated to pop-up infrastructure. With the situation rapidly developing, some of the roads shown here, such as [Kirkstall Road](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u) (in the top 10 results our methods generated for Leeds), have already been identified by Local Authorities as targets for intervention. As the recovery continues, we hope our methods will support cities to take the opportunities that the crisis presents for active travel.

<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocate to provide space for active travel, and locations of hospitals" />

<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />

With the rush to act, there is a great need for evidence of where new interventions should be prioritised. As with the medical science, research is needed now. Methods developed to identify locations of high walking and cycling potential can help ensure that the ‘pop-up’ infrastructure that goes in is effective, safe, and placed where it is most needed. Agile research can respond to rapidly changing conditions, as government priorities shift and different sectors of the economy reopen. Measures can be undertaken on a temporary basis, but if action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that they become permanent. In Paris, Covid-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).

<!-- # Where is pop-up infrastructure most needed? -->

<!-- # Story  -->

<!-- The COVID-19 pandemic has transformed all sectors of the economy. -->
<!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
<!-- Yet short-distance travel patterns have also changed greatly and, from a health and well-being perspective, this may be a more important shift. -->
<!-- With temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. -->
<!-- With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. -->
<!-- As with the medical science, research is needed now. -->
<!-- Methods developed to identify locations of high walking and cycling potential can help ensure that the 'pop-up' infrastructure that goes in now is effective, safe, and placed where it is most needed. -->

<!-- (this would be two to three sentences summarizing the article’s focus) -->

<!-- # Significance  -->

<!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> 
<!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post COVID-19 "new golden age for cycling". -->
<!-- Without evidence there is a risk that these words are not supported by effective and long-term interventions. -->
<!-- Our research into prioritising the right kind of cycling infrastructure where it is most needed can ensure that decisions made now are of maximum benefit for the coming months when physical distancing will need to continue. -->
<!-- Polls showing people do not want to 'return to normal' also suggest that the COVID-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
<!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. -->

<!-- # Key points -->

<!-- - Our analysis of data generated for the Department for Transport funded Propensity to Cycle Tool (PCT) and Cycling Infrastructure Prioritisation Toolkit (CyIPT) has already flagged dozens of roads in 10 major cities across England where a 'spare lane' could be replaced by low-cost bidirectional temporary infrastructure.  -->
<!-- - A method highlights roads based on number of lanes, cycling potential and length. -->
<!-- - Feedback from cycle campaigners coordinated by CyclingUK suggests the results for 10 cities (London, Birmingham, Manchester, Leeds, Liverpool, Newcastle, Sheffield, Cambridge, Bristol, Leicester) match intuition and can support local knowledge and community engagement in support of temporary infrastructure. -->
<!-- - Further work, which we are discussing with the Department for Transport, will develop more sophisticated methods to provide evidence that is urgently needed to support pop-up spaces in the UK and beyond. -->

<!-- CyclingUK are putting out a press release as an exclusive to national media organisations. -->
<!-- This article will provide a more detailed view to go out the following day while interest in the topic is at a peak. -->

<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post COVID-19 recovery package in Greater Manchester. -->

<!-- ![](https://raw.githubusercontent.com/cyipt/tempCycleways/master/README_files/figure-gfm/manchester-1.png) -->

<!-- Other key points(anything additional that you want to make sure they consider/know about) Highlighting the website further etc and we would also mention the BBC exclusive here. I’ve never had a problem with the Conversation agreeing to hold off on publishing an article until an embargo or exclusive deadline has been met. However it would be good to make clear how the Conversation article will differ from the BBC coverage.   -->

 <!-- supporting the case for reclaiming road space for safe, socially distanced key worker travel. -->