### 2020-05-12 conversation.Rmd
Update conversation

---
title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
---
As cities across the world prepare to emerge from lockdown, continued social distancing measures will mean that public transport options are likely to run at a fraction of their intended capacity. This is a crucial moment for urban cycling.

There is a window of opportunity to change urban travel. Acting fast, [before roads become busy again](https://www.forbes.com/sites/carltonreid/2020/05/07/car-traffic-rise-of-11-shows-urgent-need-for-pop-ups-to-protect-cyclists-and-pedestrians/amp/), will enable new and healthy active transport habits to become embedded long-term.

Our research at the University of Leeds is investigating how best to make rapid changes to urban cycling infrastructure. We have developed [methods](https://osf.io/7wjb6) to identify roads where there is space for pop-up infrastructure, and where it is most needed.

The UK government has announced £2 billion of investment in measures to promote walking and cycling in England over the next 5 years. £250 million has been allocated for "emergency interventions to make cycling and walking safer" as [Transport Secretary Grant Shapps put it](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020), with similar commitments being made by [Scottish](https://www.transport.gov.scot/news/10-million-to-support-pop-up-active-travel-infrastructure/) and [Welsh](https://gov.wales/councils-challenged-change-wales-transport-environment-during-coronavirus-lockdown) governments.

<tc-read-more id="137909"></tc-read-more>

Local authorities urgently need to decide how this funding should be spent, and where the reallocation of road space should be prioritised. New cycling infrastructure is more likely to be effective when it is developed based on analysis of the best available data, in combination with local knowledge and community engagement.

<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />

If action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that these pop-up developments can become permanent. In Paris, COVID-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).

We have experience of building tools and software to help highlight roads that may be suitable candidates for pop-up infrastructure. The [Propensity to Cycle Tool](https://www.pct.bike/), built in collaboration with other universities and funded by the Department for Transport, helps local authorities identify [where cycle networks](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) can be established.

Now, we have used this knowledge to generate results that can support investment in pop-up infrastructure. As with the PCT, the visualised and interpreted using interactive maps. The accessible and open nature of the results encourages community engagement and online debate, based on the evidence.

<tweet url="https://twitter.com/robinlovelace/status/1258411769682366466" />

In streets that are wide or have multiple lanes, road space can be reallocated to widen pavements or introduce pop-up cycleways, such as outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625). This is simple to implement,  with relatively few implications for traffic flow.

[Publicly available and interactive preliminary results](https://cyipt.github.io/popupCycleways/) for 10 cities highlight the potential of the methods and provide an opportunity for people with local knowledge to feedback, via ['issues' on the collaboration platform GitHub](https://github.com/cyipt/popupCycleways/issues) (requires a [GitHub](https://github.com/) account). Our work has already been [used](https://www.cyclinguk.org/press-release/100-miles-pop-cycle-lanes-could-enable-millions-cycle-safely-post-lockdown) to support cycling advocates to push pop-up infrastructure where it is most needed. An [interactive map of local interventions nationwide](https://www.cyclinguk.org/covid-19-safe-space-social-distancing) highlights what is possible and provides a basis for engaging with local authorities to encourage swift and effective action to encourage walking and cycling. Our research will contribute to the effort by strengthening the case for pop-up cycleways and highlighting available road space and locations with high long-term cycling potential.

The results are based on two main inputs: the Propensity to Cycle Tools's analysis of Census data under the UK Government’s [national target](https://www.gov.uk/government/publications/cycling-and-walking-investment-strategy) of doubling cycling levels by 2025, and data on the characteristics of road sections from the CyIPT project. Roads with spare lanes were defined as roads on which there are at least two lanes going in the same direction. The methods, based on open data, open source software and reproducible code, could scale nationwide and to other cities worldwide. Further refinements of the methods could include the prioritisation of roads to 'key destinations' such as hospitals (with hospitals in Leeds shown in the map below) and the inclusion of other relevant variables such as road width and safety statistics.

<image id="334113" align="centre" caption="Roads in Leeds with at least one spare lane that could be reallocated to provide space for active travel, and locations of hospitals" />

The use of open data and publicly accessible models is a key aspect of this research. Given the global nature of the challenges we face, open and collaborative research is vital so that methods can be applied by as wide a user base as possible. The potential for international application can be seen for example in [research for the World Health Organisation](https://atfutures.github.io/upthat/articles/adaptation.html) modelling cycling uptake in cities in low income countries. Within the UK, [over 50 Local Authorities](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf) are now using the PCT to inform their strategic cycle network plans.

An important [finding](https://osf.io/7wjb6) is that many cities have excess road space, under reduced traffic conditions, on routes leading to key destinations. Identifying such roads, like [Kirkstall Road](https://www.openstreetmap.org/way/7991370#map=15/53.8049/-1.5718), highlighted by the results shown in the map above and already part of plans by [Leeds City Council](https://news.leeds.gov.uk/news/new-measures-to-support-social-distancing-for-people-walking-and-cycling-in-leeds) to become a [pilot pop-up cycleway](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u), could be the key to effective investment in pop-up infrastructure. The evidence generated by our methods, in combination with community engagement, can ensure that transport interventions made during the crisis are of maximum benefit now and in the post pandemic world.

<!-- # Where is pop-up infrastructure most needed? -->

<!-- # Story  -->

<!-- The COVID-19 pandemic has transformed all sectors of the economy. -->
<!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
<!-- Yet short-distance travel patterns have also changed greatly and, from a health and well-being perspective, this may be a more important shift. -->
<!-- With temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. -->
<!-- With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. -->
<!-- As with the medical science, research is needed now. -->
<!-- Methods developed to identify locations of high walking and cycling potential can help ensure that the 'pop-up' infrastructure that goes in now is effective, safe, and placed where it is most needed. -->

<!-- (this would be two to three sentences summarizing the article’s focus) -->

<!-- # Significance  -->

<!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> 
<!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post COVID-19 "new golden age for cycling". -->
<!-- Without evidence there is a risk that these words are not supported by effective and long-term interventions. -->
<!-- Our research into prioritising the right kind of cycling infrastructure where it is most needed can ensure that decisions made now are of maximum benefit for the coming months when physical distancing will need to continue. -->
<!-- Polls showing people do not want to 'return to normal' also suggest that the COVID-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
<!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. -->

<!-- # Key points -->

<!-- - Our analysis of data generated for the Department for Transport funded Propensity to Cycle Tool (PCT) and Cycling Infrastructure Prioritisation Toolkit (CyIPT) has already flagged dozens of roads in 10 major cities across England where a 'spare lane' could be replaced by low-cost bidirectional temporary infrastructure.  -->
<!-- - A method highlights roads based on number of lanes, cycling potential and length. -->
<!-- - Feedback from cycle campaigners coordinated by CyclingUK suggests the results for 10 cities (London, Birmingham, Manchester, Leeds, Liverpool, Newcastle, Sheffield, Cambridge, Bristol, Leicester) match intuition and can support local knowledge and community engagement in support of temporary infrastructure. -->
<!-- - Further work, which we are discussing with the Department for Transport, will develop more sophisticated methods to provide evidence that is urgently needed to support pop-up spaces in the UK and beyond. -->

<!-- CyclingUK are putting out a press release as an exclusive to national media organisations. -->
<!-- This article will provide a more detailed view to go out the following day while interest in the topic is at a peak. -->

<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post COVID-19 recovery package in Greater Manchester. -->

<!-- ![](https://raw.githubusercontent.com/cyipt/tempCycleways/master/README_files/figure-gfm/manchester-1.png) -->

<!-- Other key points(anything additional that you want to make sure they consider/know about) Highlighting the website further etc and we would also mention the BBC exclusive here. I’ve never had a problem with the Conversation agreeing to hold off on publishing an article until an embargo or exclusive deadline has been met. However it would be good to make clear how the Conversation article will differ from the BBC coverage.   -->

 <!-- supporting the case for reclaiming road space for safe, socially distanced key worker travel. -->