### 2020-05-13 cities/bristol.Rmd
Update build code

# Bristol

```{r Bristol}
setwd("..")
region_name = "Bristol"
source("code/build.R", local = TRUE)
m
knitr::kable(res_table, caption = "The top 10 candidate roads for space reallocation for pop-up active transport infrastructure according to methods developed for the project.", digits = 0)
```