### 2020-05-18 code/mcr-safe-streets-analysis.Rmd
Add mcr analysis doc

---
title: "Analysis of cycling potential around key corridors in Greater Manchester"
author: "Robin Lovelace"
date: "18/05/2020"
# output: bookdown::html_document2
output: bookdown::word_document2
---

```{r, echo=FALSE, eval=FALSE}

```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(tmap)
z = pct::get_pct_zones("greater-manchester")
lads = z %>% 
  sf::st_buffer(0.0001) %>% 
  group_by(lad_name) %>% 
  summarise(all = sum(all))
# plot(lads)
l = pct::get_pct_lines("greater-manchester")
sum(z$bicycle) / sum(z$all)
sum(z$dutch_slc) / sum(z$all)
z$walk_2011 = z$foot / z$all * 100
z$cycle_2011 = z$bicycle / z$all * 100
z$pt_2011 = (z$bus + z$train_tube) / z$all * 100
z$cycle_dutch = z$dutch_slc / z$all * 100
```

```{r}
# od analysis
od = pct::get_od()
od$pt = od$train + od$light_rail + od$bus
od_mcr = od %>% filter(geo_code1 %in% z$geo_code)
# sum(od$pt) / sum(od$all)
# 0.159002
# sum(od_mcr$pt) / sum(od_mcr$all)
# 0.1434537
```

```{r}
z$active_2011 = (z$foot + z$bicycle) / z$all * 100
z$active_dutch = (z$foot + z$dutch_slc) / z$all * 100
b = 0:8 * 10
```

# Context

## Public transport

Greater Manchester has high cycling potential and high reliance on public transport (PT) .
Across the region overall, 14.3% of commutes were made by public transport in 2011, although the reliance on public transport varies substantially across the region.
Around a third of trips are made by PT in many parts of Manchester and Salford, as shown in Figure \@ref(fig:pt).
These areas areas that are particularly vulnerable to reduced capacity by public transport also tend to be within cycling distance from the main employment centre in Manchester's city centre.

```{r pt, fig.cap="The percentage of trips made by public transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("pt_2011", palette = "Blues", title = "% PT (2011)") +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

## Active transport

The geographic distribution of active transport (walking and cycling) for commuting is shown in Figure \@ref(fig:active).
13% of single mode commutes were made by active modes in 2011.
This has the potential to more than double to 33% for the region overall under the Go Dutch scenario, represented in Figure \@ref(fig:activedutch).

```{r active, fig.cap="Percentage of commutes in Greater Manchester by active transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("active_2011", palette = "Greens", title = "% Active (2011)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
# (sum(z$foot) + sum(z$bicycle)) / sum(z$all)
# (sum(z$foot) + sum(z$dutch_slc)) / sum(z$all)
```

The majority of potential growth in active transport comes from cycling due to the increased range of commutes by this mode, around 5 km for most people and further for many, due to high fitness and use of e-bikes.

```{r activedutch, fig.cap="Percentage of commutes in Greater Manchester by active transport under the Go Dutch scenario. Source: Propensity to Cycle Tool"}
tm_shape(z) + tm_fill("active_dutch", palette = "Greens", title = "% Active (Go Dutch)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

Like surrounding regions, only 2% of single mode commutes in were made by bike according to data from the 2011 Census (this figure will be higher, perhaps doubling since the introduction of pro-cycling policies since 2011 and the uptake of cycling in response to PT capacity restrictions).

Under the Go Dutch scenario, this rises to 22%, compared with 20% in Lancashire and Cheshire, with a geographic distribution shown in Figure \@ref(fig:cycling).
The region has high cycling potential because of the high population density in the region and relatively short average trip distances.

```{r cycling, fig.cap="Mode share of cycling from the 2011 Census and under the Go Dutch scenario in the Propensity to Cycle Tool", out.width="50%", fig.show='hold'}
knitr::include_graphics(c(
  "https://user-images.githubusercontent.com/1825120/82181124-d65d8600-98d9-11ea-8835-98647333ffe2.png",
  "https://user-images.githubusercontent.com/1825120/82181210-fee58000-98d9-11ea-8d9c-cc716556f769.png"
))
```

Data on walking, cycling and PT levels from the 2011 Census for MSOA zones in the region are summarised by local authority in the table below.

```{r}
table1::table1(~ walk_2011 + cycle_2011 + pt_2011 | lad_name, data = z, transpose = TRUE)
```

# Coridor analysis

# Cycling potential on key corridors