### 2020-05-18 code/mcr-safe-streets-analysis.Rmd
Update mcr analysis

---
title: "Analysis of cycling potential around key corridors in Greater Manchester"
author: "Robin Lovelace"
date: "18/05/2020"
# output: bookdown::html_document2
output: bookdown::word_document2
# output: bookdown::pdf_document2
---

```{r, echo=FALSE, eval=FALSE}

```

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, cache = TRUE, message = FALSE, warning = FALSE)
region_name = "Manchester"
library(tidyverse)
library(tmap)
z = pct::get_pct_zones("greater-manchester")
lads = z %>% 
  sf::st_buffer(0.0001) %>% 
  group_by(lad_name) %>% 
  summarise(all = sum(all))
# plot(lads)
l = pct::get_pct_lines("greater-manchester")
sum(z$bicycle) / sum(z$all)
sum(z$dutch_slc) / sum(z$all)
z$walk_2011 = z$foot / z$all * 100
z$cycle_2011 = z$bicycle / z$all * 100
z$pt_2011 = (z$bus + z$train_tube) / z$all * 100
z$cycle_dutch = z$dutch_slc / z$all * 100
```

```{r}
# od analysis
od = pct::get_od()
od$pt = od$train + od$light_rail + od$bus
od_mcr = od %>% filter(geo_code1 %in% z$geo_code)
# sum(od$pt) / sum(od$all)
# 0.159002
# sum(od_mcr$pt) / sum(od_mcr$all)
# 0.1434537
```

```{r}
z$active_2011 = (z$foot + z$bicycle) / z$all * 100
z$active_dutch = (z$foot + z$dutch_slc) / z$all * 100
z$active_ebike = (z$foot + z$ebike_slc) / z$all * 100
b = 0:8 * 10
```

# Context

## Public transport

Greater Manchester has high cycling potential and high reliance on public transport (PT) .
Across the region overall, 14.3% of commutes were made by public transport in 2011, although the reliance on public transport varies substantially across the region.
Around a third of trips are made by PT in many parts of Manchester and Salford, as shown in Figure \@ref(fig:pt).
These areas areas that are particularly vulnerable to reduced capacity by public transport also tend to be within cycling distance from the main employment centre in Manchester's city centre.

```{r pt, fig.cap="The percentage of trips made by public transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("pt_2011", palette = "Blues", title = "% PT (2011)") +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

## Active transport

The geographic distribution of active transport (walking and cycling) for commuting is shown in Figure \@ref(fig:active).
13% of single mode commutes were made by active modes in 2011.
This has the potential to more than double to 33% for the region overall under the Ebikes scenario, represented in Figure \@ref(fig:activedutch).

```{r active, fig.cap="Percentage of commutes in Greater Manchester by active transport. Source: 2011 Census"}
tm_shape(z) + tm_fill("active_2011", palette = "Greens", title = "% Active (2011)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
# (sum(z$foot) + sum(z$bicycle)) / sum(z$all)
# (sum(z$foot) + sum(z$dutch_slc)) / sum(z$all)
```

The majority of potential growth in active transport comes from cycling due to the increased range of commutes by this mode, around 5 km for most people and further for many, due to high fitness and use of e-bikes.

```{r activedutch, fig.cap="Percentage of commutes in Greater Manchester by active transport under the Ebikes scenario. Source: Propensity to Cycle Tool"}
tm_shape(z) + tm_fill("active_dutch", palette = "Greens", title = "% Active (Ebikes)", breaks = b) +
  tm_shape(lads) + tm_borders(col = "black") + tm_text("lad_name") +
  tm_scale_bar()
```

Like surrounding regions, only 2% of single mode commutes in were made by bike according to data from the 2011 Census (this figure will be higher, perhaps doubling since the introduction of pro-cycling policies since 2011 and the uptake of cycling in response to PT capacity restrictions).

Under the Ebikes scenario, this rises to 22%, compared with 20% in Lancashire and Cheshire, with a geographic distribution shown in Figure \@ref(fig:cycling).
The region has high cycling potential because of the high population density in the region and relatively short average trip distances.

```{r cycling, fig.cap="Mode share of cycling from the 2011 Census and under the Ebikes scenario in the Propensity to Cycle Tool", out.width="50%", fig.show='hold'}
knitr::include_graphics(c(
  "https://user-images.githubusercontent.com/1825120/82181124-d65d8600-98d9-11ea-8835-98647333ffe2.png",
  "https://user-images.githubusercontent.com/1825120/82255290-4e649400-994c-11ea-98c4-267fe936ccdd.png"
))
```

Data on walking, cycling and PT levels from the 2011 Census for MSOA zones in the region are summarised by local authority in the table below.

```{r}
# table1::table1(~ walk_2011 + cycle_2011 + pt_2011 | lad_name, data = z, transpose = TRUE)
# knitr::include_graphics("mcr-census-results.png")
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/82254618-10b33b80-994b-11ea-85ef-5edee3a53ac5.png")
```

# Corridor analysis

We constrained the analysis to 10 key corridors.
These provide access to the main employment hub from the North, East, South and West, as illustrated in Figure \@ref(fig:corridors).

```{r, fig.cap="Key corridors analysed for the project."}
# Airport
# Altrincham
# Leigh
# Oldham

city_centre = tmaptools::geocode_OSM(region_name, as.sf = TRUE)
city_centre_buffer = stplanr::geo_buffer(city_centre, dist = 2000)
city_centre_buffer8 = stplanr::geo_buffer(city_centre, dist = 8000)
city_centre_buffer_square = stplanr::geo_bb(scale_factor = 1.1, city_centre_buffer8)
r_pct_lanes_all = readRDS("mcr_r_pct_lanes_all.Rds")
# mapview::mapview(r_pct_lanes_all)
corridor_names = c(
  "Chester Road", # (Trafford)
  "A6",
  "A580", #A6/A580 – East Lancs to Chapel Street (Salford)
  "A34", # – Anson road/Birchfield Road (Stockport)
  "A57", # Hyde Road (Tameside)
  "A62", # A62 – Oldham Road (Oldham)
  "A56", # A56 – Bury New Road (Bury)
  "Chorlton Road", # (all Manchester)
  "A5103" # Princess road (all Manchester)
)
princess_road = r_pct_lanes_all %>% filter(name == "Princess Road")
# table(princess_road$ref) # A5103
bury = r_pct_lanes_all %>% filter(name == "Bury New Road")
# mapview::mapview(princess_road)

# load all roads using code in "code/get-data-db.R":
rsf_mcr = readRDS("mcr_rsf.Rds")
regexclude_corridors = "A537|A5149|A5102|A5034|A556|5014|5067"

key_corridors_mcr = rsf_mcr %>% 
  filter(name %in% corridor_names | ref %in% corridor_names) %>% 
  filter(str_detect(string = ref, pattern = "A")) %>% 
  filter(!str_detect(string = ref, pattern = regexclude_corridors)) %>% 
  filter(!is.na(width)) %>% 
  mutate(cycling_potential = as.numeric(pctebike)) %>% 
  select(name, ref, cycling_potential, maxspeed, aadt, width, length, idGlobal)
princess_road = key_corridors_mcr %>% filter(name == "Princess Road")

# Split up A6 and A56
# table(key_corridors_mcr$name[key_corridors_mcr$ref == "A6"])
zbs = zonebuilder::zb_zone(x = city_centre, n_circles = 1, n_segments = 4, distance = 50, segment_center = TRUE)
# plot(zbs$geometry)
# plot(princess_road, add = T)
# plot(key_corridors_mcr %>% filter(ref == "A6"), add = T)
# plot(key_corridors_mcr %>% filter(ref == "A56"), add = T)
zbs_north = zbs %>% filter(segment_id == 1)
zbs_east = zbs %>% filter(segment_id == 4)
# plot(zbs_east)
a56_north = key_corridors_mcr %>% filter(ref == "A56") %>% .[zbs_north, ]
a6_east = key_corridors_mcr %>% filter(ref == "A6") %>% .[zbs_east, ]
a57_east = key_corridors_mcr %>% filter(ref == "A57") %>% .[zbs_east, ]

key_corridors_mcr$ref[key_corridors_mcr$ref == "A56"] = "A56 (South)"
key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a56_north$idGlobal] = "A56 (North)"

key_corridors_mcr$ref[key_corridors_mcr$ref == "A6"] = "A6 (West)"
key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a6_east$idGlobal] = "A6 (East)"#

key_corridors_mcr$ref[key_corridors_mcr$ref == "A57"] = "A57 (West)"
key_corridors_mcr$ref[key_corridors_mcr$idGlobal %in% a57_east$idGlobal] = "A57 (East)"

key_corridors_mcr_grouped = key_corridors_mcr %>% 
  # group_by(name, ref) %>%
  group_by(ref) %>%
  summarise(
    name = first(name),
    cycling_potential = round(mean(cycling_potential)),
    width = round(mean(width)),
    length = round(sum(length))
    ) %>% 
  ungroup()

m = tm_shape(key_corridors_mcr_grouped) +
  tm_lines(lwd = "width", scale = 9, alpha = 0.8, col = "ref") +
  tm_shape(lads) + tm_borders() +
  tm_scale_bar()
# m
```

```{r corridors, fig.cap="Key corridors in Greater Manchester by road reference and direction relative to Manchester city centre."}
knitr::include_graphics("https://user-images.githubusercontent.com/1825120/82258452-caada600-9951-11ea-8af0-520c8407928a.png")
```

To further focus the analysis, only the sections within 8 km of Manchester city centre were included.
The results of clipping key corridors to an 8 km buffer surrounding the city centre are shown in the Figure \@ref(fig:buff8).

```{r buff8, fig.cap="Illustration of the key corridors within an 8 km buffer of Manchester city centre, with the width of road line segments proportional to estimated carriageway width. Source: www.cyipt.bike."}
key_corridors8 = key_corridors_mcr[city_centre_buffer8, ]
lads8 = sf::st_intersection(lads, city_centre_buffer_square)

m = tm_shape(key_corridors8) +
  tm_lines(lwd = "width", scale = 9, alpha = 0.8, col = "ref") +
  tm_shape(lads8) + tm_borders() + tm_text("lad_name") +
  tm_shape(city_centre_buffer8) + tm_borders() +
  tm_scale_bar() +
  tm_layout(legend.outside = TRUE)
m
```

# Cycling potential on key corridors

The cycling potential under the E-bike scenario was assessed as follows:

```{r}
key_corridors_mcr_grouped2 = key_corridors8 %>% 
  # group_by(name, ref) %>%
  group_by(ref) %>%
  summarise(
    name = first(name),
    cycling_potential = round(mean(cycling_potential)),
    width = round(mean(width)),
    length = round(sum(length))
    ) %>% 
  ungroup()

key_corridors_mcr_grouped2 %>% 
  sf::st_drop_geometry() %>% 
  group_by(ref) %>% 
  summarise(
    `Length (m)` = round(sum(length)),
    `Cycling potential` = mean(cycling_potential),
    `Width (m)` = mean(width),
    `Km/day` = round(`Length (m)` * `Cycling potential` / 1000) * 2
    ) %>% 
  arrange(desc(`Km/day`)) %>% 
  filter(`Length (m)` > 2000) %>% 
  knitr::kable(caption = "Average cycling potential, width and potential distance cycled per day for commuting (to and from work on a typical work day) for each corridor under the Ebikes scenario.")
```

Many of the key corridors have spare lanes during times of low traffic, as shown in Figure \@ref(fig:nlanes), which also shows the variability of cycling potential in segments along the different corridors.

```{r nlanes, fig.cap="Illustration of spare lanes on key corridors"}
# find the r_pct_lanes in the key roads
key_corridors_buffer = stplanr::geo_buffer(key_corridors8, dist = 20)
r_pct_lanes_in_corridors = r_pct_lanes_all[key_corridors_buffer, , op = sf::st_within]
m = tm_shape(key_corridors8) +
  tm_lines(lwd = "cycling_potential", scale = 9, alpha = 0.8, col = "ref") +
  tm_shape(r_pct_lanes_in_corridors) + tm_lines() +
  tm_scale_bar()
m
```

```{r}
l = readRDS("mcr-l-pct.Rds")
r = readRDS("mcr-r-pct.Rds")

# potential analysis ------------------------------------------------------

r$pt_mode_shift = r$ebike_sip * -1
# summary(r$pt_mode_shift)
# nrow(r) # 16363
# rnet_pt = stplanr::overline(r, "pt_mode_shift")
# saveRDS(rnet_pt, "mcr-rnet_pt.Rds")
# mapview::mapview(rnet_pt)
rnet_pt = readRDS("mcr-rnet_pt.Rds")

# since 2011 there has been a doubling

# high percentage of key workers - stats

```

Cycling potential can also be estimated in terms of the number of new cycling trips that could result from a mode shift to cycling.
Results of the shift from PT under the E-bikes scenario, which were calculated at the route segment level for the first time in this study, are shown for each key corridor in Figure \@ref(fig:shift):

```{r shift, fig.cap="Cycling potential associated with a PT/cycling mode shift under the Ebikes scenario at the route network level within 500 m of each corridor, under the (unmet) assumption that every mode has equal chance of being replaced by cycling for a given origin-destination pair. In practice, the probability of a trip being replaced by cycling is mode and context-dependent, with physical distancing rules making a PT/cycling switch proportionally more likely."}
key_corridors_buffer_100 = stplanr::geo_buffer(key_corridors8, dist = 500)
# slow operation:
rnet_pt_in_corridors = rnet_pt[key_corridors_buffer_100, , op = sf::st_within]
# saveRDS(rnet_pt_in_corridors, "rnet_pt_in_corridors.Rds")
# rnet_pt_in_corridors = readRDS("rnet_pt_in_corridors.Rds")
plot(rnet_pt_in_corridors)
# mapview::mapview(rnet_pt_in_corridors)
```

The results from this dataset will be used to assess cycling potential from a PT perspective, with results shown below.

```{r}
key_corridors_pt = st_join(rnet_pt_in_corridors, key_corridors_buffer["ref"])
key_corridors_pt$km_day = key_corridors_pt$pt_mode_shift * as.numeric(sf::st_length(key_corridors_pt)) / 1000
res_pt_corridors = key_corridors_pt %>% 
  sf::st_drop_geometry() %>% 
  group_by(ref) %>% 
  summarise(
    `Km/day` = sum(km_day)
  )
res_pt_corridors %>%
  filter(!is.na(ref)) %>% 
  arrange(desc(`Km/day`)) %>% 
  knitr::kable(digits = 0, caption = "Ranking of corridors according to estimated distance of PT trips that switch to cycling within a 500m buffer of each corridor.")
# key_corridors_pt = aggregate(rnet_pt_in_corridors, key_corridors_buffer, op = sf::st_within, FUN = mean)
# key_corridors_pt %>% 
#   sf::st_drop_geometry() %>% 
#   arrange(desc(pt_mode_shift)) 
```