### 2020-05-28 code/region_index.Rmd
Add per region landing page

---
output: html_document
title: '`r region$Name`'
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r pkg, include=FALSE}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

```{r}
knitr::include_url("m.html")
```

See full map [here](m.html).

```{r}
DT::datatable(res_table)
```