### 2020-06-05 code/region_index.Rmd
Update build + ranking

---
output: html_document
title: '`r region$Name`'
---

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r pkg, include=FALSE}
library(sf)
library(tidyverse)
library(tmap)
tmap_mode("view")
```

The map shows three key layers:

1. The **Existing cycleways** layer provides an approximation of where cycling infrastructure exists **currently** (based on OpenStreetMap data downloaded in May 2020) and is intended to help identify gaps in the network that could be filled with new interventions.
2. The **Top ranked new cycleways** layer is the central result of the analysis, providing a list of roads that have high cycling potential, a minimum threshold length, and spare space according to our definition. These may be strong candidates for reallocation of road space **immediately or in the near future** to create new cycleways on strategic corridors under the Emergency Active Travel Fund.
3. The **Cohesive network** layer is intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems. Composed of roads that have high cycling potential on some or all of their length, the layer is designed to guide **long term planning**, alongside pre-existing plans.

To see additional layers showing road sections with 'spare lanes' and width greater than 10 m, click on the layers icon in the top left, just below the +/- buttons on the map.

```{r, out.width="100%"}
knitr::include_url("m.html")
```

See full map [here](m.html), and a more description of the layers and the methodology at [cyipt.bike/rapid](https://www.cyipt.bike/rapid/)

The table below shows the top `r nrow(res_table)` roads, in terms of cycling potential and spare space criteria outlined on the landing page.
The length refers to the continuous length of road.
Cycling potential represents the average *combined number of commute and school* cycle trips that would use the road each morning under the Government Target scenario in the PCT.
'Length * potential' is the length of the road multiplied by its average cycling potential.

```{r}
DT::datatable(res_table)
```