### 2020-06-12 conversation.Rmd
Add conversation draft

---
title: "Cities must act to secure the future of urban cycling: our research shows how"
# title: "Pop-up cycleways provide a peak into post COVID-19 transport futures"
output: github_document
---

Cities worldwide are preparing for the long transition out of [lockdown](https://www.bsg.ox.ac.uk/research/research-projects/coronavirus-government-response-tracker). Physical distancing measures will be in place for many months, with impacts on all walks of life, not least transport.

With public transport options running at low capacity and emerging evidence of the role of [air quality](https://www.sciencedirect.com/science/article/pii/S0048969720321215) and [exercise](https://www.ncbi.nlm.nih.gov/pmc/articles/PMC7118448/) in mitigating the risks of COVID-19, solutions are needed more than ever.

New cycleways are being introduced in [many cities](https://twitter.com/robinlovelace/status/1268834853472088066), allowing healthy habits started during the lockdown to continue. Transport authorities must act fast, however, to take advantage of the current [cycling boom](https://www.bbc.co.uk/news/business-52564351) while [traffic levels are still below normal](https://www.forbes.com/sites/carltonreid/2020/05/07/car-traffic-rise-of-11-shows-urgent-need-for-pop-ups-to-protect-cyclists-and-pedestrians/amp/), and to avoid [gridlock](https://static1.squarespace.com/static/5d30896202a18c0001b49180/t/5ee090e5635f3e0513867b40/1591775464939/Report_1+-+Congestion+ahead_+a+faster+route+is+now+available.pdf).

<image id="341146" align="centre" caption="New cycleway on space reclaimed from motor traffic as part of the COVID-19 response in Park Lane, London." source="Transport for London." />

Our research at the University of Leeds has identified roads where there is both space and demand for cycling infrastructure.

Our [methods](https://osf.io/7wjb6) have been used in a nationwide project funded by the [Department for Transport](https://www.gov.uk/government/organisations/department-for-transport) and transport charity [Sustrans](https://www.sustrans.org.uk/) to help relieve immediate pressures on the transport system and create long term change. The result is the [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/). This is a free and open tool to help ensure that the government's [Emergency Active Travel Fund](https://www.gov.uk/government/publications/emergency-active-travel-fund-local-transport-authority-allocations) is spent where it is most needed, for maximum long term benefit.

## New routes

The [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/) provides an interactive map for every transport authority in England. The main result is a list of road sections that have both high demand and sufficient space for cycling.

These are represented as blue lines in the map of Leeds below, many of which are included in the council's [new cycleway plans](https://news.leeds.gov.uk/news/leeds-city-council-announces-emergency-walking-and-cycling-plans-in-response-to-covid-19). Additional layers in the map include existing off road cycleways (shown in green), which in many cities are disjointed or of variable quality, and, in purple, a vision of what a joined-up cycle network could look like.

<image id="340974" align="centre" source="University of Leeds" caption="The map output of the Rapid Cycleway Prioritisation Tool for West Yorkshire, zoomed in on Leeds" />

This new tool builds on [our previous work](https://www.jtlu.org/index.php/jtlu/article/view/862), carried out with other universities, to help authorities identify and develop [strategic cycle networks](https://npct.github.io/pct-shiny/regions_www/www/static/03d_other_reports/2019-use-of-pct-report.pdf). We created a national dataset of roads, including estimates of the number of lanes in either direction and road width.

This can help identify where roads have space that could be reallocated to widen pavements or to rapidly introduce new cycleways. One of the first of has been installed outside a hospital in [Leicester](https://www.leicestermercury.co.uk/news/leicester-news/light-lockdown-traffic-leicester-allows-4083625). Community engagement through projects such as [Widen My Path](https://www.widenmypath.com/), which provides a forum for comments on where more space for walking and cycling is most needed, is another vital part of this process.

<tc-read-more id="135248"></tc-read-more>

The tool's focus on road space reallocation is due to the importance of speed and capacity. Light segregation measures such as plastic bollards or [wands](https://www.bbc.co.uk/news/uk-england-leeds-52577554) can be implemented much faster than constructing new cycleways from scratch. Wide, direct and continuous cycleways, of the type created by road space reallocation on wide roads, are needed for capacity and to ensure physical distancing guidelines can be followed. Finally, road space reallocation has environmental benefits, representing '[zero carbon infrastructure](https://www.ice.org.uk/news-and-insight/the-infrastructure-blog/october-2019/net-zero-needs-zero-carbon-infrastructure)'.

An important [finding](https://osf.io/7wjb6) is that many cities have wide roads with spare space, as shown the maps of six major cities below. Notably, none of these yet have a joined-up cycle network. An example of the type of road highlighted by our tool is [Kirkstall Road](https://www.openstreetmap.org/way/7991370#map=15/53.8049/-1.5718) in Leeds, which has high cycling potential and sufficient width for new cycleways. Kirkstall road is already part of plans by [Leeds City Council](https://news.leeds.gov.uk/news/new-measures-to-support-social-distancing-for-people-walking-and-cycling-in-leeds) to become a [trial cycleway](https://www.transportxtra.com/publications/local-transport-today/news/65382/leeds-to-pilot-enhanced-cycling-safety-u).

Our methods, based on open data and code, could be used in cities worldwide. Given the global nature of the challenges we face, open and collaborative research is vital. The potential for international application can be seen in [research we carried out](https://atfutures.github.io/upthat/articles/adaptation.html) for the World Health Organisation on possible cycling uptake in cities in low income countries.

<image id="341363" align="centre" source="Robin Lovelace" caption="Maps showing existing, disjointed cycleway networks (green), potential cycleway routes on wide roads according to the Rapid Cycleway Prioritisation Tool (blue) and cohesive networks (purple) in 6 major cities." zoomable="true" />

## Planning for the future

The UK government has announced £2 billion of investment in measures to promote walking and cycling in England over the next five years. £250 million has been allocated for emergency interventions [to make cycling and walking safer](https://www.gov.uk/government/speeches/transport-secretarys-statement-on-coronavirus-covid-19-9-may-2020).

Similar commitments are being made in [Scotland](https://www.transport.gov.scot/news/10-million-to-support-pop-up-active-travel-infrastructure/) and [Wales](https://gov.wales/councils-challenged-change-wales-transport-environment-during-coronavirus-lockdown). Local authorities urgently need to decide how this funding should be spent.

<tweet url="https://twitter.com/laura_laker/status/1257660564098162689" />

If action is prioritised in places where there is a long-term need for cycle improvements, there is a greater chance that these developments can become permanent. In Paris, COVID-19 related measures aim to contribute towards Mayor Anne Hidalgo's long-term [Plan Velo](https://www.curbed.com/2020/1/15/21065343/bike-paris-cycling-anne-hidalgo).

<tc-read-more id="137909"></tc-read-more>

New cycling infrastructure is more likely to be effective when it is developed based on analysis of the best available data, in combination with vital local knowledge. City planners, politicians and citizens need to act to ensure that transport interventions made during the crisis are of maximum benefit now and in the post-pandemic world.

<!-- # Where is pop-up infrastructure most needed? -->

<!-- # Story  -->

<!-- The COVID-19 pandemic has transformed all sectors of the economy. -->
<!-- Transport patterns have been dramatically altered, with much discussion of plummeting demand for long distance travel. -->
<!-- Yet short-distance travel patterns have also changed greatly and, from a health and well-being perspective, this may be a more important shift. -->
<!-- With temporary infrastructure being put in place in dozens of cities worldwide, from [Park Lane in London](https://www.forbes.com/sites/carltonreid/2020/05/06/pop-up-cycleway-for-park-lane-as-tenfold-increase-in-cycling-predicted-for-london/#48aad0705b1e) to [Martin Luther King Jr Drive in Philadelphia](https://www.theguardian.com/world/2020/apr/11/world-cities-turn-their-streets-over-to-walkers-and-cyclists), [Auckland](https://www.roadtraffic-technology.com/news/nz-govt-extra-wide-footpaths/) to [Mexico City](https://bicitekas.org/sites/default/files/public/manuales/propuesta_ciclovias_temporales_cdmxvfinal.pdf), never has there been such a need for affordable, healthy and socially isolated transport that does not create the fumes that emerging evidence links to coronavirus susceptibility. -->
<!-- With the rush to act, there is a great need for evidence of *where* new interventions should be prioritised. -->
<!-- As with the medical science, research is needed now. -->
<!-- Methods developed to identify locations of high walking and cycling potential can help ensure that the 'pop-up' infrastructure that goes in now is effective, safe, and placed where it is most needed. -->

<!-- (this would be two to three sentences summarizing the article’s focus) -->

<!-- # Significance  -->

<!-- (why is this story interesting/relevant? Particularly to general audience, why should they care?)  --> 
<!-- Politicians of all stripes recognise the importance of acting now to prevent a surge in car use and resulting air pollution when the lockdown is gradually released. -->
<!-- Sadiq Kahn has advocated re-allocating road space for walking and cycling, and Boris Johnson has [talked](https://twitter.com/SomethingELLIS/status/1258031387334053888) of a post COVID-19 "new golden age for cycling". -->
<!-- Without evidence there is a risk that these words are not supported by effective and long-term interventions. -->
<!-- Our research into prioritising the right kind of cycling infrastructure where it is most needed can ensure that decisions made now are of maximum benefit for the coming months when physical distancing will need to continue. -->
<!-- Polls showing people do not want to 'return to normal' also suggest that the COVID-19 crisis represents an historic opportunity to readjust transport cities in favour of health. -->
<!-- If pop-up cycleways go where there is *long term need* there is a greater chance that people will demand they become permanent, as happened with the experimental cycleway in Torrington Place in London, where a 2 way road for motor vehicles has been converted into a thriving 2 way cycleway, enabling social distancing and zero emissions transport for keyworkers in central London. -->

<!-- # Key points -->

<!-- - Our analysis of data generated for the Department for Transport funded Propensity to Cycle Tool (PCT) and Cycling Infrastructure Prioritisation Toolkit (CyIPT) has already flagged dozens of roads in 10 major cities across England where a 'spare lane' could be replaced by low-cost bidirectional temporary infrastructure.  -->
<!-- - A method highlights roads based on number of lanes, cycling potential and length. -->
<!-- - Feedback from cycle campaigners coordinated by CyclingUK suggests the results for 10 cities (London, Birmingham, Manchester, Leeds, Liverpool, Newcastle, Sheffield, Cambridge, Bristol, Leicester) match intuition and can support local knowledge and community engagement in support of temporary infrastructure. -->
<!-- - Further work, which we are discussing with the Department for Transport, will develop more sophisticated methods to provide evidence that is urgently needed to support pop-up spaces in the UK and beyond. -->

<!-- CyclingUK are putting out a press release as an exclusive to national media organisations. -->
<!-- This article will provide a more detailed view to go out the following day while interest in the topic is at a peak. -->

<!-- Results for Manchester shown below, which highlights Priness Street, one of the roads where [space for active travel will be increased](https://www.manchestereveningnews.co.uk/news/greater-manchester-news/part-deansgate-close-cars-city-18178175) as part of the post COVID-19 recovery package in Greater Manchester. -->

<!-- ![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/README_files/figure-gfm/manchester-1.png) -->

<!-- Other key points(anything additional that you want to make sure they consider/know about) Highlighting the website further etc and we would also mention the BBC exclusive here. I’ve never had a problem with the Conversation agreeing to hold off on publishing an article until an embargo or exclusive deadline has been met. However it would be good to make clear how the Conversation article will differ from the BBC coverage.   -->

 <!-- supporting the case for reclaiming road space for safe, socially distanced key worker travel. -->