### 2020-06-15 rapid-slides.Rmd
Add existing cycleways slide

---
title: "The Rapid Cycleway Prioritisation Tool"
subtitle: "🛣🚲🚀<br/>methods and results<br/>[cyipt.bike/rapid](https://www.cyipt.bike/rapid/)"
author: "Robin Lovelace and Joey Talbot"
institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport). [Source code](https://github.com/cyipt/popupCycleways/blob/master/rapid-slides.Rmd)."
date: "Green Transport Recovery Series, 2020-06-17"
output:
  xaringan::moon_reader:
    css: [default, robot, robot-fonts]
    lib_dir: libs
    nature: 
      beforeInit: "https://platform.twitter.com/widgets.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, include=FALSE}
library(dplyr)
```

background-image: url(https://pbs.twimg.com/media/EZvQNJlWoAE4D_c?format=jpg&name=4096x4096)
background-position: bottom
background-size: contain

# Motivations

--

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

1) reduced capacity on public transport 2) need for space so physical distancing guidelines can be followed
3) healthy restart 4) need to act fast

---

### Existing cycleways

- Aims to identify **immediate** issues
- From OpenStreetMap, March 2020
- Contains sections of variable quality
- Aim to highligh where gaps need to be filled

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cycleways-leeds.png)

---

### How to identify ‘top ranked new cycleways’?

<!-- ![](figures/results-top-leeds.png) -->

- Aims to identify **short term priorities**
- All road sections with spare lanes or estimated width > 10 m (CyIPT)
- Of those, which have a minum continuous length of at least 500 m?
- Rank the roads by cycling potential (commute + school, Government Target)
- Show the top n, where n depends on the size of the region (70 in London)

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-top-leeds.png)

---

Tablular results for Leeds

```{r, echo=FALSE, out.height="100%"}
res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
res_table = res_leeds %>%
  sf::st_drop_geometry() %>% 
  arrange(desc(mean_cycling_potential))
# %>% 
#   select(
#     Name = name,
#     Reference = ref,
#     `Length (m)` = group_length,
#     `Cycling potential` = mean_cycling_potential,
#     `Length * potential (km)` = km_cycled,
#     # `SRN` = srn,
#     `Speed limit` = speed_limit,
#     `Mean estimated width (m)` = mean_width,
#     `Majority spare lane(s)?` = majority_spare_lane
#     ) 
DT::datatable(res_table, height = 600)
```

---

### How to identify a ‘cohesive network’?

- Aims to support **long term planning**
- Intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems
- Find  roads that have high cycling potential on some or all of their length
- Join them up

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cohesive-leeds.png)

---

# How can I find the results for my area?

---

# How should I use these results?

---

background-image: url("https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png")
background-position: right
background-size: contain

--

Every city in England

## Results

---

# Case study: Leeds

.left-column[

Results in Leeds:

```{r, echo=FALSE, results='markup'}
res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
res_leeds %>% 
  arrange(desc(mean_cycling_potential)) %>% 
  slice(1:5) %>%
  sf::st_drop_geometry() %>% 
  dplyr::select(name) %>% 
  as.data.frame()
```

Cycleway planned for

Kirkstall Road (source: [BBC](https://www.bbc.co.uk/news/uk-england-leeds-52577554))

]

.right-column[

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)

See the results at [www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire)

]

---

---

## What are the key limitations of the tool?

---

## Where can I find further information about the tool?

---

## How to cope with reduced public transport capacity post COVID-19?

Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways:

![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)

---

# Data

---

# Methods

---

# Results

---

# Demo

---

---

# Links I

- Sheffield [infrastructure](https://www.cyclesheffield.org.uk/2018/01/01/cyclesheffield-response-to-portobello-trippet-lane-cycle-route-designs/)

- Article on Yorkshire Voice on the Leeds-Bradford [Cycle Superhighway](http://www.yorkshire-voice.com/mixed-views-as-29m-cycle-superhighway-linking-leeds-and-bradford-nears-completion/)

- A detailed article on the same cycleway by [Mark Strong](https://aseasyasridingabike.wordpress.com/2016/08/11/a-visit-to-the-leeds-bradford-cycle-superhighway/)

- Academic article on thinking at level of cycleway networks (Buehler and Dill [2015](https://www.tandfonline.com/doi/full/10.1080/01441647.2015.1069908?journalCode=ttrv20))

- The Propensity to Cycle Tool ([pct.bike](https://www.pct.bike/))

- The Cycling Infrastructure Prioritisation Toolkit ([CyIPT.bike](https://www.cyipt.bike/existing/#13/53.4690/-2.2506/mapnik))

- Consultation on the Torrington Place / Tavistock Place 'pop-up' cycleway from [Cambden.gov.uk](https://www.camden.gov.uk/documents/20142/3452947/Consultation+leaflet+FINAL.pdf/f628d6e8-c47b-82f1-cb40-8e24db78bea6)

---

# Links II

- Map of pop-up interventions in response to COVID-19 from [CyclingUK](https://www.cyclinguk.org/covid-19-safe-space-social-distancing)

- Maps of roads with 'spare lanes' in 10 cities from our our 'cyipt' [website](https://www.cyclinguk.org/article/10-cities-where-pop-bike-lanes-could-benefit-millions-explore-maps)

- New statutory [guidance from DfT on road space reallocation](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities)

- Analysis to support [road space reallocation in Manchester](https://github.com/cyipt/popupCycleways/blob/master/code/mcr-safe-streets-analysis.md#analysis-of-cycling-potential-around-key-corridors-in-greater-manchester)

- Image of new [pop-up going in in Greater Manchester](https://github.com/cyipt/popupCycleways/issues/27)

---

# Limitations and next steps

- Source: Adam Reynolds on a [GitHub issue](https://github.com/luukvdmeer/sfnetworks/issues/13)

![](https://user-images.githubusercontent.com/1948692/74239070-0429a200-4ccf-11ea-8ce0-5530e84db2ff.jpeg)

---

# Thanks!

Contact: me at

- r. lovelace at leeds ac dot uk (email)

- [`@robinlovelace`](https://twitter.com/robinlovelace) (Twitter)

--

- Check out www.pct.bike

--

- www.cyipt.bike

--

- Pop-up cycleway info: watch this space

```{r, echo=FALSE, eval=FALSE}
# export code from slides
knitr::purl("tds-regional-street.Rmd")
```