### 2020-06-15 rapid-slides.Rmd
Update slides, close cyipt/icicle#18

---
title: "The Rapid Cycleway Prioritisation Tool"
subtitle: "methods and results<br/>[cyipt.bike/rapid](https://www.cyipt.bike/rapid/)"
author: "Robin Lovelace and Joey Talbot"
institute: "University of Leeds, [Institute for Transport Studies](https://environment.leeds.ac.uk/transport).<br/>Publicly available [slides](https://www.cyipt.bike/rapid/slides/). Reproducible [code](https://github.com/cyipt/popupCycleways/blob/master/rapid-slides.Rmd). Paper [preprint](https://osf.io/7wjb6/)."
date: "Green Transport Recovery Series, 2020-06-17"
output:
  xaringan::moon_reader:
    css: [default, robot, robot-fonts]
    lib_dir: libs
    nature: 
      beforeInit: "https://platform.twitter.com/widgets.js"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r, engine='bash', eval=FALSE, echo=FALSE}
# 🛣🚲🚀
cp rapid-slides.html popupCycleways/v1/slides/index.html
cp -Rv libs popupCycleways/v1/slides/
cd popupCycleways
git status
git add -A
git commit -am 'Add slides'
git push
cd -
```

```{r, eval=FALSE, echo=FALSE}
# Aim: make pdf slides
pagedown::chrome_print("rapid-slides.html")
browseURL("rapid-slides.pdf")
piggyback::pb_upload("rapid-slides.pdf")
piggyback::pb_download_url("rapid-slides.pdf")
```

```{r, include=FALSE}
library(dplyr)
```

background-image: url(https://pbs.twimg.com/media/EZvQNJlWoAE4D_c?format=jpg&name=4096x4096)
background-position: bottom
background-size: contain

# Motivations

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

1) reduced capacity on public transport 2) need for space so physical distancing guidelines can be followed
3) healthy restart 4) need to act fast. Image credit: [TfL](https://twitter.com/W_Bradley/status/1268836932219478016).

---

Plans from Leeds City Council responding to national [guidance](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities) and [funding](https://www.gov.uk/government/news/2-billion-package-to-create-new-era-for-cycling-and-walking) for 'pop-up' cycleways (image credit: [Leeds City Council](https://news.leeds.gov.uk/news/leeds-city-council-announces-emergency-walking-and-cycling-plans-in-response-to-covid-19)):

![](https://pbs.twimg.com/media/EZ_-A0dXgAAlBzt?format=png&name=900x900)

---

### Existing cycleways

- Aims to identify **immediate** issues
- From OpenStreetMap, March 2020
- Contains sections of variable quality
- Aim to highligh where gaps need to be filled

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cycleways-leeds.png)
Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)

---

### How to identify ‘top ranked new cycleways’?

<!-- ![](figures/results-top-leeds.png) -->

- Aims to identify **short term priorities**
- All road sections with spare lanes or estimated width > 10 m (CyIPT)
- Of those, which have a minum continuous length of at least 500 m?
- Rank the roads by cycling potential (commute + school, Government Target)
- Show the top n, where n depends on the size of the region (70 in London)

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-top-leeds.png)

Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)

---

Tablular results for Leeds

```{r, echo=FALSE, out.height="100%"}
res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
res_table = res_leeds %>%
  sf::st_drop_geometry() %>% 
  arrange(desc(mean_cycling_potential))
# %>% 
#   select(
#     Name = name,
#     Reference = ref,
#     `Length (m)` = group_length,
#     `Cycling potential` = mean_cycling_potential,
#     `Length * potential (km)` = km_cycled,
#     # `SRN` = srn,
#     `Speed limit` = speed_limit,
#     `Mean estimated width (m)` = mean_width,
#     `Majority spare lane(s)?` = majority_spare_lane
#     ) 
DT::datatable(res_table, height = 600)
```

---

### How to identify a ‘cohesive network’?

- Aims to support **long term planning**
- Intended to show what a joined-up cycle network could look like if we were to consider new cycleways by either closing roads to motorised traffic or creating one-way systems
- Find  roads that have high cycling potential on some or all of their length
- Join them up

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-cohesive-leeds.png)
Image credit: [Rapid Cycleway Prioritisation Tool](https://www.cyipt.bike/rapid/)

---

### How can I find the results for my area?

.left-column[

Head to [cyipt.bike/rapid/](https://www.cyipt.bike/rapid/)

]

.right-column[

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/results-regions.png)
]

---

# How should I use these results?

- Aim: to support evidence-based prioritisation of bids for Tranche 2 of the Emergency Active Travel Fund
- Complementing local knowledge and stakeholder engagement
- Scoping of options
- Order of implementations

Investment should depend on many factors including:

- cycling potential (default metric to rank schemes)
- total length of cycling infrastructure created
- cycling potential multiplied by road section length

---

background-image: url("https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/facet-output.png")
background-position: right
background-size: contain

--

Every city in England

## Results

---

# Case study demo: Leeds

.left-column[

Results in Leeds:

```{r, echo=FALSE, results='markup'}
res_leeds = sf::read_sf("https://www.cyipt.bike/rapid/west-yorkshire/top_routes.geojson")
res_leeds %>% 
  arrange(desc(mean_cycling_potential)) %>% 
  slice(1:5) %>%
  sf::st_drop_geometry() %>% 
  dplyr::select(name) %>% 
  as.data.frame()
```

Cycleway planned for

Kirkstall Road (source: [BBC](https://www.bbc.co.uk/news/uk-england-leeds-52577554))

]

.right-column[

![](https://raw.githubusercontent.com/cyipt/popupCycleways/master/figures/mapout.png)

See the results at [www.cyipt.bike/rapid/west-yorkshire](https://www.cyipt.bike/rapid/west-yorkshire)

]

---

## Limitations

- Narrow focus on road space reallocation options
- No economic appraisal
  - Use the Active Mode Appraisal Toolkit ([AMAT](https://www.gov.uk/government/publications/tag-social-and-distributional-impacts-worksheets)) 
- The best locations for new cycleways may not always have space
  - Search for parallel routes, which may be available
- Many options not in top routes
  - see cohesive network and look for other option
- The road network data is from 2017 and widths are estimates
  - Compare with real world knowledge
- Only work and school trips included
  - Use local knowledge of new developments and other 'trip attractors
- Prioritises routes in large cities in authorities with multiple settlements
  - Use alternative sources of information, including the PCT
  - Do additional analysis, e.g. based on data downloaded from the Rapid tool or the PCT

---

## Where can I find further information about the tool?

- See the technical specification [here](https://github.com/cyipt/popupCycleways/releases/download/v1/report.pdf)

Data about the tools informing the model can be found here:

- The [Propensity to Cycle Tool](https://www.pct.bike/)
- The [Cycling Infrastructure Prioritisation Tool](https://www.cyipt.bike)
- [OpenStreetMap](https://www.openstreetmap.org/#map=7/52.486/-2.428&layers=C)

- Code underlying the project can be accessed [here](https://github.com/cyipt/popupCycleways).
- Support the project with technical input via the [GitHub issue tracker](https://github.com/cyipt/popupCycleways/issues/14)
- Data generated by the project can be downloaded from GitHub (e.g. see GeoJSON files for London [here](https://github.com/cyipt/popupCycleways/tree/gh-pages/v1/greater-london)).

---

# Links

- Academic article on thinking at level of cycleway networks (Buehler and Dill [2015](https://www.tandfonline.com/doi/full/10.1080/01441647.2015.1069908?journalCode=ttrv20))

- The Propensity to Cycle Tool ([pct.bike](https://www.pct.bike/))

- The Cycling Infrastructure Prioritisation Toolkit ([CyIPT.bike](https://www.cyipt.bike/existing/#13/53.4690/-2.2506/mapnik))

- Consultation on the Torrington Place / Tavistock Place 'pop-up' cycleway from [Cambden.gov.uk](https://www.camden.gov.uk/documents/20142/3452947/Consultation+leaflet+FINAL.pdf/f628d6e8-c47b-82f1-cb40-8e24db78bea6)

- Map of pop-up interventions in response to COVID-19 from [CyclingUK](https://www.cyclinguk.org/covid-19-safe-space-social-distancing)

- New statutory [guidance from DfT on road space reallocation](https://www.gov.uk/government/publications/reallocating-road-space-in-response-to-covid-19-statutory-guidance-for-local-authorities)

- Analysis to support [road space reallocation in Manchester](https://github.com/cyipt/popupCycleways/blob/master/code/mcr-safe-streets-analysis.md#analysis-of-cycling-potential-around-key-corridors-in-greater-manchester)

- Image of new [pop-up going in in Greater Manchester](https://github.com/cyipt/popupCycleways/issues/27)

---

## Next steps

Many future directions of travel including evidence to support [Low Traffic Neighbourhoods](https://github.com/luukvdmeer/sfnetworks/issues/13)

![](https://user-images.githubusercontent.com/1948692/74239070-0429a200-4ccf-11ea-8ce0-5530e84db2ff.jpeg)

---

# Thanks!

Contact: me at

- r. lovelace at leeds ac dot uk (email)

- [`@robinlovelace`](https://twitter.com/robinlovelace) (Twitter)

--

- Check out www.pct.bike

--

- www.cyipt.bike/rapid

<!-- Tweet: @ launch -->

<!-- The Rapid Cycleway Prioritisation Tool is finally launched 🚀 -->

<!-- It provides evidence to support road space reallocation for active travel in response to #COVID19. -->

<!-- Article: -->
<!-- Preprint of methods: -->
<!-- Web tool (free and #opensource):  -->

<!-- CWIS technical reports -->
<!-- Cost per intervention type -->